Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2016-06-28 Projects/DelCo_2012/Report/DelCo_2012.tex

For the 12 middle position replicate comparisons (Fig.~\ref{fig:CAFS_Water_mid}), the results (via the dashed red line) point to water having a slight advantage in gas cooling relative to CAFS. At the back position (Fig.~\ref{fig:CAFS_Water_full}), for the three replicate comparisons, the least squared error line falls within the experimental uncertainty. This indicates an inability to distinguish between the two suppression agents with statistical confidence. The data can be further divided to analyze the impact of the hose stream time on comparison between water and CAFS. Figures~\ref{fig:CAFS_Water_78_mid} and \ref{fig:CAFS_Water_SS_mid} show the results for 7 replicate comparisons of a 7/8~in. solid stream and 5 replicate comparisons of a straight stream respectively; both aimed at the middle position.

@@

The 7 replicate tests comparing water and CAFS from a 7/8~in. solid stream nozzle aimed at the middle position show a large bias towards water. For the 5 replicate tests with a straight stream nozzle, the line fit through the data falls right at the edge of where there is still uncertainty regarding an assessment of which suppression agent had more cooling.

The back position can also be decomposed as function of stream. Figure~\ref{fig:CAFS_Water_SS_back} shows a single comparison between water and CAFS for a straight stream in the back position while Fig.~\ref{fig:CAFS_Water_Fog_back} shows a 2 replicate comparison between water and CAFS for a 30$^{\circ}$ Fog in the back position.

@@

For the straight stream comparison (Fig.~\ref{fig:CAFS_Water_SS_back}), there is an inability to distinguish between the suppression agents with statistical confidence. The 30$^{\circ}$ Fog replicate comparison (Fig.~\ref{fig:CAFS_Water_Fog_back})

@@

@@

>>>>

For the 12 middle position replicate comparisons (Fig.~\ref{fig:CAFS_Water_mid}), the results (via the dashed red line) show that water has a slight advantage in gas cooling relative to CAFS. At the back position (Fig.~\ref{fig:CAFS_Water_full}), for the 3 replicate comparison, the least squared error line falls within the experimental uncertainty. This indicates an inability to distinguish between the two suppression agents with statistical confidence. The data can be further divided to analyze the impact of the nozzle type on comparison between water and CAFS. Figures~\ref{fig:CAFS_Water_78_mid} and \ref{fig:CAFS_Water_SS_mid} show the results for 7 replicate comparisons of a 7/8~in. solid stream and 5 replicate comparisons of a straight stream respectively; both aimed at the middle position.

@@

The 7 replicate tests comparing water and CAFS from a 7/8~in. solid stream nozzle aimed at the middle position show a large bias towards water. For the 5 replicate tests with a straight stream nozzle, the least squared error line fit through the data falls along the edge of the measurement uncertainty bounds on the water side of comparison.

The back position can also be decomposed to compare stream type. Figure~\ref{fig:CAFS_Water_SS_back} shows a single comparison between water and CAFS for a straight stream in the back position while Fig.~\ref{fig:CAFS_Water_Fog_back} shows a 2 replicate comparison between water and CAFS for a 30$^{\circ}$ Fog in the back position.

@@

For the straight stream comparison (Fig.~\ref{fig:CAFS_Water_SS_back}), there is an inability to distinguish between the suppression agents with statistical confidence. The 30$^{\circ}$ Fog replicate comparison (Fig.~\ref{fig:CAFS_Water_Fog_back}) shows that CAFS has better cooling potential compared to water.

@@

To better visualize the data presented in Table~\ref{tab:Test_Results}, Fig.~\ref{fig:fs_1_1} shows the thermocouple temperature and heat flux changes for CAFS Test 1 as a result of the hallway (first) and room (second) suppression actions.

\begin{figure}[ht!]
	\includegraphics[width=\columnwidth]{../Figures/Pictures/Metric/DelCoFogTest1FirstSuppression}
	\includegraphics[width=\columnwidth]{../Figures/Pictures/Metric/DelCoFogTest1SecondSuppression}
	\caption{Impact of fire suppression actions on thermocouple and heat flux measurements for CAFS Test 1. The top image shows the hallway suppression impact and the bottom images shows the room suppression impact.}
	\label{fig:fs_1_1}
\end{figure}

@@

\chapter{Fire Suppression Figures}
\label{app:fire_suppression}

\begin{figure}[ht!]
	\includegraphics[width=\columnwidth]{../Figures/Pictures/Metric/DelCoFogTest1FirstSuppression}
	\includegraphics[width=\columnwidth]{../Figures/Pictures/Metric/DelCoFogTest1SecondSuppression}
	\caption{Impact of fire suppression actions on thermocouple and heat flux measurements for CAFS Test 1. The top image shows the hallway suppression impact and the bottom images shows the room suppression impact.}
	\label{fig:app_fs_1}
\end{figure}

\begin{figure}[ht!]
	\includegraphics[width=\columnwidth]{../Figures/Pictures/Metric/DelCoFogTest2FirstSuppression}
	\includegraphics[width=\columnwidth]{../Figures/Pictures/Metric/DelCoFogTest2SecondSuppression}
	\caption{Impact of fire suppression actions on thermocouple and heat flux measurements for Water Test 1. The top image shows the hallway suppression impact and the bottom images shows the room suppression impact.}
	\label{fig:app_fs_2}
\end{figure}

### 2016-06-28 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex

\Title{Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard}
\Author{Daniel Madrzykowski}
\Year{2015}

@@

\abstract{
Uncontrolled fires result in loss of life, property and can create an adverse economic impact on a community.  The investigation of fires provides a means to identify the cause of the fire in order to develop a knowledge base that could enable the elimination of that cause and thus reduce the losses from fires.  However, as the result of the review of several arson homicide cases and a review forensics by the U.S. National Academy of Sciences, questions about the lack of the use of science in the practice of fire investigation were raised.  Specifically the National Academy of Sciences indicated that ``... research is needed on the natural variability of burn patterns...''
This study addresses that need in two ways:
\begin{enumerate}
\item Examining the repeatability of several small fire sources and the fire patterns that were generated by those fires
\item Examining the capability of using numerical models to simulate the fires and the resulting fire patterns based on input data collected from engineering reference sources, and bench scale and full scale fire experiments conducted in this study.
\end{enumerate}
This paper highlights the many uncertainties involved in what appear to be simple fire experiments.  A variety of uncertainties related to the composition of the fuels, measurement methods, and analysis techniques just to list a few of the areas considered.  The objective of the research was to examine the repeatability of pre-flashover fire patterns generated from exposure to short duration (300 s maximum) well characterized fires.  Three different fuels were used: natural gas, gasoline, and polyurethane foam.  Each fuel had a similar top surface area.  The heat release rate data showed that the variability was greater for more complex fuels.  While the variation in peak heat release rate with the natural gas was similar to the expanded uncertainty of the measurement system, 11~\%, the variation in the peak heat release rates of the gasoline and the polyurethane foam increased due to increased uncertainties in the burning behavior of the fuels.  The patterns generated by the polyurethane foam fire, had more variability than the natural gas and the gasoline fire patterns.  This is likely due to the greater variability of the burning of the solid fuel and the lower heat release rate.  The maximum fire pattern heights generated from the natural gas and gasoline fires were shown to have uncertainties of 18~\% or less based on a Type A statistical analysis with 95~\% confidence limits.  The comparison of the fire pattern heights and the mean flame height demonstrated that the ``steady state'' natural gas fires exhibited the highest level of agreement and polyurethane foam fueled fire exhibited the least agreement, with the gasoline fueled fires in between.

@@

\chapter{Introduction}
\label{chapter:Introduction}

Uncontrolled fires result in loss of life, property and can create an adverse economic impact on a community.  Findings from an archaeological site in South Africa, published in 2012, indicate that mankind has productively used fire for approximately 1 million years~\cite{Berna:2012}.  This is 300 thousand years earlier than previously thought.  Despite mankind’s extensive history with fire, according to the World Health Organization, an estimated 265,000 people die each year as a result of burn injuries~\cite{WHO:2014}.  Based on an estimated global population of 7.3 billion people, the death rate due to burns from all sources would be 3.6 per 100,000 populations.  The World Fire Statistics Center collects data from European, North American and a few Asia-Oceania countries.  Within this group of countries the fire deaths per 100,000 population range from 0.02 (Singapore) to 2.03 (Finland)~\cite{Climate:2014}. There are a significant number of factors that determine the variability of fire death rates between nations, socio-economic conditions being a leading factor~\cite{WHO:2014}.

>>>>

\Title{Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard}
\Author{Daniel Madrzykowski}
\Year{2016}

@@

\abstract{
Unwanted fires result in loss of life and property.  These fires can also create an adverse economic impact on a community.  The investigation of fires provides a means to identify the cause of the fire in order to develop a knowledge base that could enable the elimination of that cause and thus reduce the losses from unwanted fires. Questions about the lack of science in the practice of fire investigation have been raised during the review of several arson homicide cases and a forensics science review by the U.S. National Academy of Sciences.  Specifically the National Academy of Sciences indicated that ``... research is needed on the natural variability of burn patterns...''
This study addresses that need in two ways:
\begin{enumerate}
\item Examining the repeatability of several small fire sources and the fire patterns that were generated by those fires
\item Examining the capability of numerical models to simulate the fires and the resulting fire patterns based on input data collected from engineering reference sources, and bench-scale and full-scale fire experiments conducted in this study.
\end{enumerate}
This paper highlights the many uncertainties involved in what appear to be simple fire experiments.  Uncertainties related to the composition of the fuels, measurement methods, and analysis techniques are a few of the areas considered.  The objective of the research was to examine the repeatability of pre-flashover fire patterns generated from exposure to short duration (300 s maximum) well characterized fires.  Three different fuels were used: natural gas, gasoline, and polyurethane foam.  Each fuel had a similar top surface area.  The heat release rate data showed that the variability was greater for more complex fuels.  The variation in peak heat release rate with the natural gas was similar to the expanded uncertainty of the measurement system, 11~\%.  However, the variation in the peak heat release rates of the gasoline and the polyurethane foam increased due to increased uncertainties in the burning behavior of the fuels.  The patterns generated by the polyurethane foam fire, had more variability than the natural gas and the gasoline fire patterns.  This is likely due to the greater variability of the burning of the solid fuel and the lower heat release rate.  The maximum fire pattern heights generated from the natural gas and gasoline fires were shown to have uncertainties of 18~\% or less based on a Type A statistical analysis with 95~\% confidence limits.  The comparison of the fire pattern heights and the mean flame height demonstrated that the ``steady state'' natural gas fires exhibited the highest level of agreement and polyurethane foam fueled fire exhibited the least agreement, with the gasoline fueled fires in between.

@@

\chapter{Introduction}
\label{chapter:Introduction}
\section{Background}

The need for fire investigation and value to society, tools used , example.

\subsection{The need and purpose for Fire Investigation}
Unwanted fires result in loss of life, property and can create an adverse economic impact on a community.  Recent findings from an archaeological site in South Africa, indicate that mankind has productively used fire for approximately 1 million years~\cite{Berna:2012}.  This is 300 thousand years earlier than previously thought.  Despite mankind's extensive history with fire, according to the World Health Organization, an estimated 265,000 people die each year as a result of burn injuries~\cite{WHO:2014}.  Based on an estimated global population of 7.3 billion people, the death rate due to burns from all sources would be 3.6 per 100,000 people.  The World Fire Statistics Center collects data from European, North American and a few Asia-Oceania countries.  Within this group of countries the fire deaths per 100,000 population range from 0.02 (Singapore) to 2.03 (Finland)~\cite{Climate:2014}. There are a significant number of factors that determine the variability of fire death rates between nations, socio-economic conditions being a leading factor~\cite{WHO:2014}.
